Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-09 paper/paper.tex
link to source code
empirical evaluation in progress

\usepackage[color]{changebar}
\usepackage{enumitem}
\usepackage[lighttt]{lmodern}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}

@@

We run this program for \texttt{n} equal to 10, 15, 20, and 25. Clojure
is consistently only twice as fast as Anglican
(Table~\ref{tbl:hanoi}); considering the amount of overhead
Anglican introduces, this is a good result.
\begin{table}
    \centering
    \begin{tabular}{r|r|r|r|r}
        {\bf n}  & {\bf 10} & {\bf 15} & {\bf 20} &{\bf 25} \\ \hline
        Clojure &  0.32 & 1.45 & 38.1  &  1245. \\ \hline
        Anglican & 0.63 & 3.00 & 75.9 & 2381. 
    \end{tabular}
    \caption{Towers of Hanoi. Running times, in
    seconds, of Clojure and Anglican, averaged over 100 runs.
    Clojure is only twice as fast as Anglican.}
    \label{tbl:hanoi}
\end{table}

The other comparison must be made between Anglican and another
probabilistic programming environment. WebPPL~\cite{GS15} is
similar to Anglican in expressivity.  WebPPL runs on top of
\href{http://v8project.blogspot.com/}{V8}, a high performance
JavaScript engine featuring a native compiler.
Table~\ref{tbl:webppl} shows results for two models from
\href{https://github.com/probmods/webppl}{WebPPL
distribution}, \textit{LDA} and
\textit{Logistic regression}. For
comparison, we re-implemented the models in Anglican, as close
as possible to the WebPPL implementations, and run the same
inference algorithm (called MCMC in WebPPL, LMH in Anglican)
with the same number of iterations. Anglican outperforms WebPPL
on both models. This can be attributed both to a better
transformed representation of probabilistic programs and to more
efficient architecture of the inference engine.
\begin{table}
    \centering
    \begin{tabular}{r|c|c}
        {\bf model}  & {\bf LDA} & {\bf Logistic regression} \\ \hline
        Anglican & 22. & 4.8 \\ \hline
        WebPPL & 31. & 9.4
    \end{tabular}

    \caption{Anglican vs. WebPPL. Running times, in seconds, of
    Anglican and WebPPL, for $100\,000$ iterations of Markov Chain
    Monte Carlo (Lighweight Metropolis-Hastings), averaged over
    100 runs. Anglican outperforms WebPPL on both models.}
    \label{tbl:webppl}
\end{table}

>>>>

\usepackage{siunitx}
\usepackage[color]{changebar}
\usepackage{enumitem}
\usepackage[lighttt]{lmodern}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}

@@

We run this program for \texttt{n} equal to 10, 15, 20, and 25. Clojure
is consistently only twice as fast as Anglican
(Table~\ref{tbl:hanoi}); considering the amount of overhead
Anglican introduces, this is a good result.
\begin{table}
    \centering
    \begin{tabular}{r|r|r|r|r}
        {\bf n}  & 10 & 15 & 20 & 25 \\ \hline
		{\bf Anglican} & 0.63 & 3.00 & 75.9 & 2381.  \\ \hline
		{\bf Clojure} &  0.32 & 1.45 & 38.1  &  1245.
    \end{tabular}
    \caption{Towers of Hanoi. Running times, in
    seconds, of Clojure and Anglican, averaged over 100 runs.
    Clojure is only twice as fast as Anglican.}
    \label{tbl:hanoi}
\end{table}

The other comparison must be made between Anglican and another
probabilistic programming environment. WebPPL~\cite{GS15} is
similar to Anglican in expressivity.  WebPPL runs on top of
\href{http://v8project.blogspot.com/}{V8}, a high performance
JavaScript engine featuring a native compiler.
Table~\ref{tbl:webppl} shows results for a few models from
\href{https://github.com/probmods/webppl}{WebPPL distribution},
as well as from Anglican test suite. For the comparison, we
re-implemented WebPPL models in Anglican, and Anglican models in
WebPPL, following as close as possible the original
implementations, and run the same inference algorithm (called
MCMC in WebPPL, LMH in Anglican) with the same number of
iterations. Anglican outperforms WebPPL on all models. This can
be attributed both to a better transformed representation of
probabilistic programs and to more efficient architecture of the
inference engine.
\begin{table}
    \centering
	\begin{tabular}{r|S[table-format=2.2]|S[table-format=2.2]}
		{\bf model}  & {\bf Anglican} & {\bf WebPPL} \\ \hline
		LDA & 22. & 31. \\ 
		\shortstack{Linear regression}  & 0. & 0. \\
		\shortstack{Logistic regression} & 6.2 & 9.4 \\
		\shortstack{Multivariate regression} & 0. & 0. \\
		\shortstack{Simple branching} & 0. & 0.
    \end{tabular}

    \caption{Anglican vs. WebPPL. Running times, in seconds, of
    Anglican and WebPPL, for $100\,000$ iterations of Markov Chain
    Monte Carlo (Lighweight Metropolis-Hastings), averaged over
    100 runs. Anglican outperforms WebPPL on all models.}
    \label{tbl:webppl}
\end{table}
The source code of the examples is available in the
\href{https://bitbucket.org/probprogs/anglican-white-paper}{paper
repository}.
